### 2024-01-24 latex_files/supplementary_material.tex

\documentclass{article}
\usepackage{booktabs} % For better quality table rules
\usepackage{caption} % For better caption format

\begin{document}

\section{List of Materials and Their Markets}

\begin{table}[ht]
\centering
\caption{A comprehensive list of various materials in the default configuration.}
\label{tab:materials}

\begin{tabular}{ll}
\toprule
\textbf{Market Name} & \textbf{Material group} \\
\midrule
market for aluminium & aluminium \\
market for antimony & antimony \\
market for bauxite & bauxite \\
market for beryllium & beryllium \\
market for bismuth & bismuth \\
market for cadmium & cadmium \\
market for calcium borates & borates \\
market for cement & cement \\
market for cerium & cerium \\
market for chromium & chromium \\
market for coal & coal \\
market for cobalt & cobalt \\
market for coke & coke \\
market for copper & copper \\
market for dysprosium & dysprosium \\
market for erbium & erbium \\
market for europium & europium \\
market for electricity & electricity \\
market for ferroniobium & niobium \\
market for fluorspar & fluorspar \\
market for gadolinium & gadolinium \\
market for gallium & gallium \\
market for gold & gold \\
market for graphite & graphite \\
market for hafnium & hafnium \\
market for helium & helium \\
market for holmium & holmium \\
market for hydrogen & hydrogen \\
market for indium & indium \\
market for latex & latex \\
market for lithium & lithium \\
market for magnesium & magnesium \\
market for natural gas & natural gas \\
market for nickel & nickel \\
market for palladium & palladium \\
market for petroleum & petroleum \\
market for phosphate & phosphate rock \\
market for platinum & platinum \\
market for rare earth & rare earth \\
market for rhodium & rhodium \\
market for sand & sand \\
market for selenium & selenium \\
market for scandium & scandium \\
market for silicon & silicon \\
market for silver & silver \\
market for sodium borates & borates \\
market for strontium & strontium \\
market for tantalum & tantalum \\
market for tellurium & tellurium \\
market for tin & tin \\
market for titanium & titanium \\
market for uranium & uranium \\
market for tungsten & tungsten \\
market for vanadium & vanadium \\
market for vegetable oil & vegetable oil \\
market for tap water & water \\
market for water & water \\
market for zinc & zinc \\
market for zirconium & zirconium \\
\bottomrule
\end{tabular}
\end{table}

\end{document}