### 2024-04-01 latex_files/main.tex

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times,authoryear]{elsarticle}
%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,3p,times,authoryear]{elsarticle}
%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,5p,times,authoryear]{elsarticle}
\documentclass[review,3p,authoryear]{elsarticle}
%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}
\usepackage{booktabs}
% ---------------------------------------------------------------------
%             FOR DRAFTING / REVIEW (DELETE BEFORE SUBMISSION)
% ---------------------------------------------------------------------
\usepackage{caption}
\usepackage{subcaption}
\usepackage[T1]{fontenc}
\usepackage[british]{babel}
\usepackage{microtype}
\usepackage[tt=false]{libertine}
\usepackage{xcolor}
\definecolor{blue}{RGB}{33,150,209}
\definecolor{red}{RGB}{255,0,0}
\usepackage{xurl}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{pdfpages}
\usepackage{enumitem}
\usepackage{lineno}
\usepackage{float}
\setlength{\linenumbersep}{5pt}
% Save the old \texttt command
\let\oldtexttt\texttt
% Redefine the \UrlFont to use a smaller typewriter font
\renewcommand{\UrlFont}{\ttfamily\small}
% Redefine the \texttt command to use a smaller font size and allow line breaks
\renewcommand{\texttt}[1]{{\ttfamily\small\nolinkurl{#1}}}

\usepackage{tcolorbox}
\newcommand{\cbox}[1]{
    \begin{tcolorbox}[hbox, colback=red!5!white, colframe=red!65!black, boxrule=0.25pt, boxsep=2pt, left=2pt, right=2pt, top=1pt, bottom=1pt]
        \small\sffamily #1
    \end{tcolorbox}
}

% ------------------------------------------------------

\journal{Resources, Conservation and Recycling}

\begin{document}

\linenumbers{}

% SPOTLIGHTS (5 bullet points, max 120 characters each)

{\Large Spotlights}
\vspace{1em}
\begin{itemize}
    % Bullet 1: Critical context and background information on the problem addressed
    \item Tracking waste and material flows is essential for developing the circular economy and supply chain resilience
    % Bullet 2: A brief overview of the key finding of the study (or findings if necessary)
    \item A tool was developed to quantify waste and material footprints of activities in life cycle assessment (LCA) databases
    % Bullet 3: The most radical, creative, disruptive or innovative aspect of the manuscript
    \item Built to complement the Brightway LCA framework and ActivityBrowser, the tool is customisable and easy to use
    % Bullet 4: The significance of the results to the environment, economics, or society
    \item The tool can be used to identify pre-consumer waste and material hotspots that are often hidden in supply chains
    % Bullet 5: Future vision or the most important implications for continued research
    \item Improved data availability and quality would enable more detailed and accurate waste and material footprinting
\end{itemize}

\newpage
%% GRAPHICAL ABSTRACT
\begin{graphicalabstract}
    \includegraphics{grabs.jpg}
\end{graphicalabstract}

%% RESEARCH HIGHLIGHTS (3-5 bullet points, max 85 characters each)
\begin{highlights}
    \item T-reX, a new software tool for quantifying waste and material flows in LCA
    \item Assesses supply risks by calculating demand for critical materials
    \item Simplifies quantification of user-specified waste and material categories
    \item Rapidly identifies waste and material demand hotspots
    \item Presents a case study of five Li-ion battery supply chains
\end{highlights}

\begin{frontmatter}

    \title{T-reX: A tool to quantify waste and material footprints in current and future Life Cycle Assessment (LCA) databases}
    \author[1]{Stewart Charles McDowall\corref{cor1}}
    \ead{s.c.mcdowall@cml.leidenuniv.nl}
    \author[1]{Elizabeth Lanphear}
    \author[1]{Stefano Cucurachi}
    \author[1]{Carlos Felipe Blanco}

    \cortext[cor1]{Corresponding author:}

    \affiliation[1]{
        organization={Institute of Environmental Sciences (CML), Leiden University},
        addressline={P.O. Box 9518},
        city={Leiden},
        postcode={2300 RA},
        state={South Holland},
        country={The Netherlands}}

    \begin{abstract}
        % \cbox{Abstract word count: 155, Limit: 150}
        The quintessential principle of the `circular economy' is to minimise material consumption and waste generation.
        Thus, identifying and quantifying waste and material flows are of fundamental importance. Life Cycle Assessment (LCA) is a powerful method for this, given its capacity to pinpoint hotspots of environmental impact throughout the life cycle of activities, where the implementation of circular principles could be most effective.

        Introducing T-reX, a Python tool extending the brightway ecosystem, allowing facile quantification of user-defined supply chain demands in current and prospective scenarios. T-reX streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.

        A case study of five lithium ion batteries demonstrates T-reX's utility, quantifying categorised waste and material inventory footprints, and thus, their potential environmental burdens. T-reX can aid sustainable decision-making and contribute to the development of the `circular economy' by facilitating analysis of material consumption and waste generation in LCA.

    \end{abstract}

    % KEYWORDS (max 6)
    \begin{keyword}
        circular economy \sep{}waste \sep{}material \sep{}life cycle assessment \sep{}critical raw material \sep{}supply chain
    \end{keyword}

\end{frontmatter}
\cbox{Total word count: ca. 5200, Limit: 5000}
\cbox{Note for reviewers: please annotate the pdf and email it to me at \href{mailto:sc.mcdowall@cml.leidenuniv.nl}{sc.mcdowall@cml.leidenuniv.nl}}
\cbox{Supplementary material is available \href{https://github.com/Stew-McD/T-reX_Publication/blob/main/reviews/T-reX_Manuscript_V3/T-reX_manuscript_V3_SupplementaryMaterial.pdf}{here}}
\cbox{Style guide for the journal: \href{https://www.sciencedirect.com/journal/resources-conservation-and-recycling/publish/guide-for-authors}{here}}

% list of abbreviations
\begin{table}[H]
    \centering
    \caption{List of abbreviations}\label{tab:abbreviations}
    \begin{tabular}{ll}
        \toprule
        CF            & Carbon Footprint                                                    \\
        EF            & Ecological Footprint                                                \\
        CML           & Centrum voor Milieuwetenschappen (Centre for Environmental Science) \\
        CRM           & Critical Raw Material                                               \\
        CPC           & Cooperative Patent Classification                                   \\
        CSI           & Crustal Scarcity Indicator                                          \\
        CSP           & Crustal Scarcity Potential                                          \\
        EDIP          & Environmental Design of Industrial Products                         \\
        EEIOA         & Environmentally Extended Input-Output Analysis                      \\
        EOL           & End-of-Life                                                         \\
        FOEN          & Swiss Federal Office for the Environment                            \\
        IAM           & Integrated Assessment Model                                         \\
        IE            & Industrial Ecology                                                  \\
        LCA           & Life Cycle Assessment                                               \\
        LCI           & Life Cycle Inventory                                                \\
        LCIA          & Life Cycle Impact Assessment                                        \\
        LFP           & Lithium Iron Phosphate                                              \\
        Li-ion        & Lithium ion                                                         \\
        LiMn\(_2\)O\(_4\) & Lithium Manganese Oxide                                             \\
        MF            & Material Footprint                                                  \\
        MFA           & Material Flow Analysis                                              \\
        NCA           & Nickel Cobalt Aluminum                                              \\
        NMC           & Nickel Manganese Cobalt                                             \\
        PWF           & Product Waste Footprint                                             \\
        RCP           & Representative Concentration Pathway                                \\
        ReCiPe        & A standard LCIA method set                                          \\
        SSPs          & Shared Socioeconomic Pathways                                       \\
        T-reX         & Tool for re-X in LCA                                                \\
        UBP           & Umweltbelastungspünkte (Environmental Impact Points)                \\
        UN            & United Nations                                                      \\
        WF            & Waste Footprint                                                     \\
        \bottomrule
    \end{tabular}
\end{table}

%% main text
\section{Introduction}\label{sec:introduction}
%\cbox{Section word count: 2000}
\input{introduction.tex}

\section{Methodology}\label{sec:methodology}
%\cbox{Section word count: 1700}
\input{methodology.tex}

\section{Results}\label{sec:results}
% \cbox{Section word count: 1200}
\input{results.tex}

\section{Discussion}\label{sec:discussion}
% \cbox{Section word count: 450}
\input{discussion.tex}

\section{Conclusions}\label{sec:conclusions}
% \cbox{Section word count: 150}

We have created `T-reX', an extension to the \texttt{brightway} ecosystem that enables users to calculate the waste and material footprints of a product or service in an LCA database. It explodes the database, identifies upstream waste and material exchanges, edits them, and writes matching custom `pseudo' LCIA methods. These exchanges become pseudo-biosphere flows and thus, the footprint can be calculated as with the existing LCIA methods. T-reX can be easily customised by the user to calculate the footprints of other supply chain flows such as water, gas, and critical raw materials.

This work extends the state of knowledge by exploring the relationship between various waste aggregation methods and environmental damage indicators, contributing to a deeper understanding of life cycle waste inventories and their association with supply chain risk and potential environmental damage.

% Back matter
\section*{Data availability}
All data used in this analysis are publicly available online under the noted sources.T-reX is installable via the Python Package Index (PyPI) and is available at \url{https://pypi.org/project/T_reX_LCA}.
The full source code for T-reX is available at \url{https://www.github.com/Stew-McD/T-reX}.
A user guide and comprehensive documentation are available at \url{https://T-reX.readthedocs.io}.

% \section*{CRediT authorship contribution statement}
% \cbox{Co-authors, please check this and change as necessary.}
% \textbf{Stewart Charles McDowall:} Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing - original draft, Writing - review \& editing, Visualization.
% \textbf{Elizabeth Lanphear:} Conceptualization, Methodology, Software, Validation, Writing - review \& editing.
% \textbf{Stefano Cucurachi:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.
% \textbf{Carlos Felipe Blanco:} Conceptualization, Writing - review \& editing, Supervision, Project administration, Funding acquisition.

% \subsection*{Alternative CRediT statement}
\section*{CRediT authorship contribution statement}

\begin{itemize}[noitemsep]
    \item \textbf{SCM}: Stewart Charles McDowall
    \item \textbf{EL}: Elizabeth Lanphear
    \item \textbf{SC}: Stefano Cucurachi
    \item \textbf{CFB}: Carlos Felipe Blanco
\end{itemize}

\includegraphics[width=\columnwidth, height=7cm, keepaspectratio]{credit.pdf}

\section*{Declaration of competing interest}
The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

\section*{Acknowledgements}
Part of this research project was financially supported by the European Union's Horizon 2020 research and innovation programme under the grant agreement No. 101058522 (project FutuRaM --- futuram.eu). The authors would like to thank the reviewers for their valuable comments and suggestions.

The authors would like to acknowledge those who were part of the futures thinking small group session at the FutuRaM consortium meeting in Berlin in June 2023 in which the term `T-reX/T(ool)-reX' was coined to name the vision presented by Stewart Charles McDowall of a comprehensive opensource software framework for the study and development of secondary raw material systems. The authors hope to continue the development of T-reX toward this vision.

\section*{Supplementary material}\label{sec:supplementary}
The supplementary material contains the following:
\begin{enumerate}
    \item Metadata of the T-reX Python package
    \item Details of the modules and user configuration
    \item Description of the computational workflow
    \item Modular flowchart of T-reX
    \item Example of the terminal output of T-reX
    \item Complete tabulated results of the case study
    \item Complete visualisations of the case study
\end{enumerate}

\bibliographystyle{elsarticle-harv}
\bibliography{references.bib}

% \appendix

% \section{Example output}
% \label{app:example_output}

\end{document}