### 2024-04-08 latex_files/discussion.tex
word_slash_LL

Waste generation and material demand are strongly associated with the environmental impacts of human activities~\citep{laurenti2023wastefootprint,steinmann2017resourcefootprints, demirer2019wastefootprint}, and thus must be represented in LCA accounting. Although there are existing LCIA methods that provide endpoint impact scores related to material demand and waste generation, they contain convoluted formulae or subjective weighting, or their complexity and lack of transparency can make them difficult to use and interpret~\citep{foen2021ecofactors,hauschild2003edip,cen2019en15804, arvidsson2020csi,foen2021ecofactors}.

T-reX advances the state-of-the-art in LCA by providing practitioners with a simple, flexible, and transparent way to calculate supply chain waste and material footprints, delivering results in standard units and as direct aggregations of the relevant demand inventories. Once LCA databases in the \texttt{brightway} project have been processed with T-reX, the user can easily apply (and re-apply) the T-reX `pseudo-LCIA' methods to calculate the waste and material inventory footprints in the same way that they would with a conventional LCIA method.

The simple case study of five Li-ion batteries presented in this paper demonstrated the utility, flexibility, and limitations of T-reX.

By adjusting the user configuration for future scenarios and waste/material categorisation, we produced a set of customised versions of current and prospective \texttt{ecoinvent} databases along with a `pseudo-biosphere' T-reX database. Then, by applying the T-reX `pseudo-LCIA' methods, categorised waste and material inventory footprints for present and future supply chains were trivial to calculate. Additionally, visual exploration in \texttt{ActivityBrowser} was possible as the T-reX `pseudo-LCIA' methods are structurally identical to standard LCIA methods when integrated within the \texttt{brightway} project.

One limitation of T-reX is that it does not yet provide specific information in a readily accessible format on the composition of the waste generated. This information is required to thoroughly assess the potential environmental impacts of this waste. Currently, the user would need to manually explore the waste footprint inventory produced by the application of the T-reX to determine if the waste generated represents an actual loss of resources or environmental risk, or, for example, is simply a transfer of the `overburden' in mining activity, which is classified as `inert waste'. A methodical classification of waste exchanges and end-of-life fates is expected to be facilitated by the increasingly detailed and disaggregated data in each successive release of \texttt{ecoinvent}~\citep{fitzgerald2023ecoinventdocumentation}.

The utility of T-reX in studies of future supply chains is limited by the fact that the currently available prospective databases focus largely on changes in the energy, steel, cement, and transport sectors~\citep{sacchi2023premisedocs}. As demonstrated in the results of the case study---where there were often very few scenario-temporal changes in many waste and material footprint indicators---the utility of prospective LCA is restricted if there is inadequate adoption of future background inventories. In particular, the inclusion of scenarios with future waste processing technology could greatly improve our predictions of waste and material flows and offer valuable insight into their potential impacts~\citep{bisinella2024wastelca}. A strong focus on enhancing these prospective databases is, thus, of critical importance to the future of prospective LCA, and by extension, to the development of the `circular economy'.