### 2024-09-22 reviews/external/JRC-review/T-reX_response-to-review_external_JRC.md
it-builds!

Dear Reviewers and Editor,

We would like to extend our sincere thanks for the time and effort you have invested in reviewing our manuscript on T-reX. We are particularly grateful for the detailed and constructive feedback provided. We have carefully addressed each of your comments in the responses below and have made the necessary revisions to the manuscript. We hope that you find the revised version satisfactory.

### Reviewer One

#### Comment:

“I think that this manuscript needs to be ameliorated.”

**Summary:**The reviewer requested more detail for the case study, as well as a sensitivity or uncertainty analysis.Key points raised:

- Comment 1: System boundary
- Comment 2: Input-output data
- Comment 3: Material footprint data (with more discussion, and a call for an ‘overall material footprint’)
- Comment 4: Sensitivity or uncertainty analysis

#### Author Responses:

##### Comment 1.1 – System Boundary:

_“The system boundary of the case study should be defined more clearly, possibly with diagrams.”_

**Response 1.1:**

(Methodology section 2.2)
We have used the most basic system boundary for this case study, since we are mearly trying to show what the programme can do, not really to provide particular results. We have added another sentence to define the system boundaries more clearly. That being only the Li-ion battery markets, as modelled in ecoinvent version 3.9.1.

##### Comment 1.2 – Input-Output Data:

_“No input-output table provided for the case studies. Please provide this inventory data in the manuscript.”_

**Response 1.2:**
For this example, described in the Methodology section 2.2, the inventory data is taken directly from the relevant activities in the ecoinvent database (version 3.9.1). No adaption was made by the authors to this data. This has been stated more clearly in the text.

##### Comment 1.3 – Material Footprint Data:

_“Results and Discussion section: the study presented some material footprint data. I recommend further discussion and interpretation of the results. Additionally, it is necessary to present the overall material footprint of the battery.”_

**Response 1.3:**
As detailed in the Methodology (section 2.1.1 and 2.1.2) Results (section 3.2), and Discussion sections (first two paragraphs), the material footprint results generated by T-reX are aggregations of the exchanges in the model system targeted by the user’s search criteria. In this case, the default selection includes XX materials and their flows as represented in ecoinvent. The authors believe that presenting a ‘total material footprint’ based on the sum of these individual materials would be misleading, as the LCA models represent only specific aspects of the system. A ‘total material footprint’ without consideration of the material type is not meaningful. One of T-reX’s strengths is its ability to allow users to focus on materials of interest. We have added clarification in the Discussion section (lines X-Y) to make this context clearer to the reader.

##### Comment 1.4 – Sensitivity and Uncertainty Analysis:

_“The authors should also perform a sensitivity analysis or uncertainty analysis.”_

**Response 1.4:**
The battery data used in this case study is taken directly from ecoinvent activities and serves primarily as a basic example to demonstrate the programme’s capabilities. Therefore, we believe that a sensitivity or uncertainty analysis on the case study results is unnecessary and could distract from the overall message of the paper. However, to ensure clarity, we have expanded the discussion on the limitations of the case study results mainto make this point more explicit to the reader.